J-ReCoVer: Java Reducer Commutativity Verifier

### 2018-11-13 draft.tex
Trim trailing spaces in `draft.tex`
polish

MapReduce framework for data-parallel computation was first proposed by Google~\cite{dean04} and later implemented in the Apache Hadoop project~\cite{hadoop}.
Under the MapReduce framework, the reducer computes output values from a sequence of input values transmitted over the network.  Due to the non-determinism in data transmission, the order of input values arrived at the reducer is not fixed.
The \emph{commutativity problem} of reducers asks if the output of a reducer is independent of the order of its inputs. There are several advantages for a reducer being commutative, e.g., the verification problem of a MapReduce program can be reduced to the problem of verifying a sequential program. 
In this paper, we propose effective heuristics for reducer commutative analysis and implement them as a tool J-ReCoVer (Java Reducer Commutativity Verifier), which is the first tool that is specialized in checking reducer commutativity. Experimental results over 128 benchmark examples collected from literatures and open repositories are very positive; J-ReCoVer correctly handles over 97\% of them.

@@

MapReduce is one of the most popular framework for data parallel computation.
A MapReduce~\cite{dean04,hadoop} program consists of several pairs of \emph{mappers} and \emph{reducers} running on a machine cluster for handling big data in parallel. Usually mappers and reducers are the only components in a MapReduce program that involves concurrency. Mappers read data from a distributed database and output a sequence of \emph{key-value} pairs. The sequence elements (key-value pairs) with the same key will be sent to the same reducer for further processing. Due to scheduling polices and network latency, the same inputs may arrive a reducer in different order in different executions, which results in the so-called \emph{commutativity problem} of reducers~\cite{csallner13testing,xiao14mr,ChenHSW15,ChenSW16}, that is, if the outputs of a reducer is independent of the order of its inputs. 
If all reducers are commutative, they will have the same external behavior under all possible schedules. In such a case, by fixing a schedule, the verification problem of a MapReduce program can be reduced to the verification problem of a sequential program, which is known to be much easier then concurrent verification. One needs to consider only a particular interleaving of input values instead of all of them.

@@

We use two sets of integer reducer benchmarks to evaluate J-ReCoVer. The first set is collected from open repositories such as Github and Bitbuckect. With the help of a search engine~\cite{searchcode} over those repositories, we collected 118 examples. J-ReCoVer is able to correctly analyze all but 3 of them. We also tried all 10 integer reducers descried in verification literatures~\cite{ChenHSW15,ChenSW16} and J-ReCoVer can handle all of them.

\subsection*{Related Works}
Reducer commutativity problem can be reduced to program equivalence problem. One creates another program $R'$ that first non-deterministically swaps two consecutive input values and then executes the code of $R$. If $R'$ and $R$ are equivalent, using the fact that all permutations of a list can be obtained by swapping consecutive list elements finitely many times\footnote{Here is an example to produce [2;3;5;1;4] from [1;2;3;4;5] by swapping consecutive elements: [\textbf{1};\textbf{2};3;4;5] $\rightarrow$ [2;\textbf{1};\textbf{3};4;5] $\rightarrow$ [2;3;1;\textbf{4};\textbf{5}]$\rightarrow$ [2;3;\textbf{1};\textbf{5};4]$\rightarrow$ [\textbf{2};\textbf{3};5;1;4]$\rightarrow$ [2;3;5;1;4].}, $R$ can be proved to be commutative. There are a series of research results on program equivalence checking (or closely related topics such as regression verification)~\cite{symdiff,fedyukovich2015automated,sharma2013data,godlin2009regression,fedyukovich2016property,felsing2014automating,KieferKlebanovUlbrich2017,lahiri2013differential,grossman2017verifying,barthe2011relational,KlebanovRuemmerUlbrich2017}. At a high level, checking equivalence of two programs $P$ and $P'$ can be reduced to a sequential verification problem by executing $P'$ after $P$ and at the end checking if they always have the same outputs. It can be made more efficient by finding the right~\emph{synchronization points} and combining the code of $P$ and $P'$ in an interleaved manner. A lot of the research effort in this research direction is in finding good synchronization points. For the case of commutativity analysis, good synchronization points are already given; one can simply use the beginning of the loop as the synchronization point. However, in our experience, if we naively reduce the commutativity problem to equivalence problem and check it in a precise manner, many examples cannot be solved. In this paper, we propose a good over-approximation (c.f. Section~\ref{sec:prover}) of the reducer's behavior. The approximation allows a much more efficient yet precise enough commutativity analysis.

@@

\section{Notations and Definitions}
\label{section:integer-reducers}
For the presentation purpose, we describe a highly simplified model to characterize the essence of reducers. The model allows to describe the computation of reducers and investigate their commutativity.

\begin{equation*}
\begin{array}{rclcl}
\multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and Constant}\\
e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
g\in \Grd  & {\triangleq}  & e < e \mid not\ e \mid g \wedge g  \mid \ldots& & \textmd{Guard}\\
s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && \textmd{Command}\\
p\in \Prg  & {\triangleq}  & s.\rloop\{s\}.s&& \textmd{Reducer Program}\\
\end{array}
\end{equation*}

Here $v\in \Var$ and $c \in \Z$ are variables and constant integer values, respectively. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{guard} in $\Grd$ can be a comparison between two expressions or boolean combinations of several guards. A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of several commands, or $out(v)$ that outputs that value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each branch\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing all if statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g)\wedge s_1$ or $(not\ g) \wedge s_2$.} . According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. A few examples of reducer programs in this model can be found in Figure~\ref{fig:reducer_example}.

Some reducer uses two loops to compute the output. For example, for standard derivation, one first computes the average of inputs and then uses it to compute the standard derivation. In that case, we can still verify the reducer by first partitioning it into two reducers and then verifying them separately. One reducer computes the average (Figure~\ref{fig:reducer_example}(b)) and the other reducer uses a non-deterministic value to over-approximate the computed average ($a := *$) and and use it in the computation of standard derivation (Figure~\ref{fig:reducer_example}(c)). The reducer in Figure~\ref{fig:reducer_example}(c) ``over-approximates''  all possible output of the standard derivation computation; it computes an output even if $a$ is not the average of inputs. However, the reducer is still commutative and the commutativity verification can still be done.

\begin{figure}
	\begin{minipage}{0.32\textwidth}
		\begin{algorithm}[H]
			$m := \cur$; \;
			\Loop{}{
				$t:=\cur$;\;
				\uIf{ $t > m$}{
					$m := t$ 
				}
			};
			$out(m)$\;
		\end{algorithm}
		\caption*{(a) max}
	\end{minipage}	
	\hspace{-0.5cm}
	\begin{minipage}{0.3\textwidth}
		\begin{algorithm}[H]
			$s := 0;c:=0$; \;
			\Loop{}{
				$s := s+\cur$;\;
				$c := c+1$ 
			};
			$o := s/c$;\;
			$out(o)$\;
		\end{algorithm}
		\caption*{(b) average}
	\end{minipage}	
	\begin{minipage}{0.4\textwidth}
		\begin{algorithm}[H]
			$a:=*;s := 0; c:=0$;\;
			\Loop{}{
				$s := s+(\cur-a)^2$; \;
				$c := c+1$ 
			};
			$o := s/c$;\;
			$out(o)$
		\end{algorithm}
		\caption*{(c) standard derivation}
	\end{minipage}	
	\caption{Example of reducer programs}
	\label{fig:reducer_example}
\end{figure}

@@

\section{Overview of the J-ReCoVer tool}
\label{sec:overview}
%Hadoop, Java;
The input of J-ReCoVer is a reducer program written in Java, which is the most popular programming language used in the Hadoop~\cite{hadoop} MapReduce framework. 
The J-ReCoVer tool has three main components, the \emph{Preprocessor}, the \emph{Prover}, and the \emph{BugFinder}. As the name suggests, the Preprocessor reads as input a reducer program and performs the required preprocessing.  The goal of Prover is to show that a given reducer is commutative and the goal of BugFinder is the opposite. The architecture of J-ReCoVer can be found in Figure~\ref{fig:overview-jrecover}. The user can input a reducer program to J-ReCoVer either though our web-interface or binary installed in their own machine.

@@

\textbf{The Preprocessor component} first compiles a reducer program to bytecode and use the tool Soot~\cite{soot} to further convert it to the so called ``Jimple'' format, which is an intermediate language that is designed to make the analysis of Java program easier. Under the Hadoop MapReduce framework, the permutation of the input is handled by the scheduler/shuffler component and is affected by things like network latency, which is not controllable by the user. In order to deal with this issue, we wrote our own dummy Hadoop interface for the reducer as a part of the Preprocessor component. So that the input order of the reducer is now controlled by J-ReCoVer. Finally, we perform in the Preprocessor some simple program transformations tasks to simplify the analysis. For example, we convert all output commands $out(e)$ inside loops to an special assignment.

\textbf{The BugFinder component} randomly generated 1500 pairs of lists. The lists in a pair are permutations of each other. A concrete counterexample is reported if inputting any pair of lists from the 1500 generated ones to the reducer generates different results.
Our procedure for generating random pairs is quite naive. We use five different input list of lengths 5, 7, 9 ,11, and 13. For each length, we generates 300 lists. For each list, we pick uniformly at random one of its permutations. Although the approach is simple, in practice it finds counterexamples in all of our non-commutative benchmarks.

\textbf{The Prover component} reduces the commutativity problem to a SMT formula satisfiability problem, using the fact that all permutations of a list can be obtained by swapping consecutive elements finitely many times. At a high level, we check equivalence between a reducer program and its variant that have two consecutive inputs swapped. We show that this can be reduced to a SMT problem and give it to SMT solver Z3~\cite{z3} for solving. In case that Z3 proved the formula unsatisfiable, J-ReCoVer stops and reports that the reducer is commutative. 
If we run the approach in a naive manner, it does not scale to large programs and is very imprecise.  We will introduce two critical optimizations that enables the proposed approach working for real-world examples in Section~\ref{section:optimizations}: a data flow analysis to extract critical variables and an improved symbolic execution algorithm to generate Z3 formulae. More details of the algorithms used in the Preprocessor and the Prover components will be explained in Section~\ref{sec:preprocessor_prover} and the optimizations will be introduced in~\ref{section:optimizations}.

@@

\subsection{The First Program Transformation Task in the Preprocessor}
\label{sec:program_trans1}
J-ReCoVer  first generates a new variable $v_{out_n}$ for the $n$-th output command $out(e)$ inside $s_2$ and replace the $out(e)$ command with an assignment $v_{out_n}:=e$. 
Notice that the value of $v_{out_n}$ stays the same from the point it has been assigned a value. 
If the values of $v_{out_n}$ are the same for all permutations of input lists at the end of $s_2$, we know that the reducer is commutative w.r.t. to all output commands inside the loop body $s_2$.

\subsection{The Prover: Reducing Commutativity to SMT Solving}
\label{sec:prover}
Below we begin with a simple version where $s_1$ never calls the $\cur$ function. Later we will explain how to handle the general case using program transformation. The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_n): \Z^{n+1} \rightarrow \Z^n$ returns a tuple of values $x_1',x_2',\ldots,x_n'$, where $n$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively. The construction of $F$ from $s_2$ can be done in the standard way. We will introduce a more efficient data structure for the construction in Section~\ref{section:optimizations}.

We check if the formula is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_n$.
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_n)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_n) )
\label{eq:commu}
\end{equation}

@@

In the example of ``average'', the values of $s$ and $c$ are the same no matter the reducer first reads $n_1$ and then $n_2$ or reads the two values in the swapped order.
So it computes the same value for the input [1;2;3;4;5] and any other input with two consecutive values swaps, e.g., [1;2;\textbf{4};\textbf{3};5].
Since any permutation of a sequence can be obtained by swapping consecutive values finitely many times~\cite{algebra}, this further implies that starting from the same variable valuation, the loop ends with the same valuation for any permutation of input values. Recall that the first program transformation task remembers all output values in variables that will stay unchanged until the end of the loop. The validity of Equation~(\ref{eq:commu}) also implies that the execution of the loop produces the same sequence of outputs.

We explain why that the validity of Equation~(\ref{eq:commu}) is sufficient to prove the commutativity of the reducer. Notice that $s_1$ (the commands before entering the loop) does not read from the input (the assumption in this subsection). The execution of $s_1$ is deterministic in the sense that from the same initial valuation, it outputs the same sequence of values and ends at the same valuation. From the validity of Equation~(\ref{eq:commu}), any permutation of input values leads the reducer to the same final valuation and the same sequence of outputs. Before executing $s_3$ (the commands before after the execution of the loop), all inputs have been read (this is the exit condition of the loop) already, similar the the case of $s_1$, it behaves deterministically for all permutations of inputs and outputs the same sequence of values.

@@

In real-world examples, it is often that the $s_1$ part of the reducer also reads from the input. In the example in Figure~\ref{fig:reducer_max} (a), the reducer program remembers the first input value in the variable $m$, increases its value by 10, and then updates its value to bigger ones if any occurs in the loop. The corresponding Equation~(\ref{eq:commu}) is valid in this case, however, the reducer is not commutative. For example, the two lists $[1,2,3,4,5]$ and $[5,4,3,2,1]$ are permutations of each other, but the reducer respectively outputs $11$ and $15$ with them as the inputs.

We handle the problem again by program transformation. To handle the example in Figure~\ref{fig:reducer_max} (a), we move the prefix $s_1$ into the loop body and use a new variable $s$ to force that $s_1$ is always executed before the original loop body. The result after the transformation can be found in Figure~\ref{fig:reducer_max} (b). Observe that the new reducer program over-approximates all possible inputs/outputs of the original one.
It also satisfies all assumptions required for the Prover component: (a) The commands before entering the loop does not have any $\cur$ expression and (II) each branch (recall that ``branch'' is formally defined in Section~\ref{section:integer-reducers}) of the loop body has at most one $\cur$. So we can apply Equation~(\ref{eq:commu}) to prove its commutativity.
Since the new reducer has all the possible outputs of the original one, the new reducer is commutative implying that the original one is also commutative.

\begin{wrapfigure}{r}{0.56\textwidth} 
		\vspace{-1.2cm}
	\begin{algorithm}[H]
		$s'_0;s:=*;$\;
		\Loop{}{
			\uIf{$s=1$}{ $x_{i_1} := \cur; s'_1; s:=2$}
			\uElseIf{ $s =2$}{ $x_{i_2} := \cur; s'_2; s:=3$}
			\uElseIf{ $s = 3$}{$\ldots$}
			\uElseIf{ $s = m$}{$x_{i_m} := \cur; s'_m; s:=m+1$}
			\uElse{ $s_2$}
		};$s_3$\;
	\end{algorithm}
	\caption{The second transformation task}
	\label{fig:general_program_transformation}
		\vspace{-1.2cm}
\end{wrapfigure}

For the general case, we first rewrite the $s_1$ part of the reducer $s_1;\rloop\{s_2\};s_3$ to the form 
$$s'_0;x_{i_1}{:=} \cur;s'_1.\ldots x_{i_{m}}{:=}\cur;s'_m, $$ where $\cur$ never occur in $s'_0,s'_1,\ldots, s'_m$. Then we produce the reducer in Figure~\ref{fig:general_program_transformation}, which is an over-approximation of the original reducer satisfying all assumptions required for the Prover. Hence the Equation~(\ref{eq:commu}) can then be applied to prove its commutativity.

\section{Optimizations}
\label{section:optimizations}
In this section we explain how data flow analysis are used in our case to improve the precision of commutativity checking and the algorithm we used to construct the $F$ function in Equation~(\ref{eq:commu}).

>>>>

MapReduce framework for data-parallel computation was first proposed by Google~\cite{dean04} and later implemented in the Apache Hadoop project~\cite{hadoop}.
Under the MapReduce framework, the reducer computes output values from a sequence of input values transmitted over the network.  Due to the non-determinism in data transmission, the order of input values arrived at the reducer is not fixed.
The \emph{commutativity problem} of reducers asks if the output of a reducer is independent of the order of its inputs. There are several advantages for a reducer being commutative, e.g., the verification problem of a MapReduce program can be reduced to the problem of verifying a sequential program. 
In this paper, we propose effective heuristics for reducer commutative analysis and implement them as a tool J-ReCoVer (Java Reducer Commutativity Verifier), which is the first tool that is specialized in checking reducer commutativity. Experimental results over 128 benchmark examples collected from literature and open repositories are very positive; J-ReCoVer correctly handles over 97\% of them.

@@

MapReduce is one of the most popular framework for data parallel computation.
A MapReduce~\cite{dean04,hadoop} program consists of several pairs of \emph{mappers} and \emph{reducers} running on a machine cluster for handling big data in parallel. Usually mappers and reducers are the only components in a MapReduce program that involves concurrency. Mappers read data from a distributed database and output a sequence of \emph{key-value} pairs. The sequence elements (key-value pairs) with the same key will be sent to the same reducer for further processing. Due to scheduling polices and network latency, the same inputs may arrive at a reducer in different order in different executions, which results in the so-called \emph{commutativity problem} of reducers~\cite{csallner13testing,xiao14mr,ChenHSW15,ChenSW16}, that is, if the outputs of a reducer is independent of the order of its inputs. 
If all reducers are commutative, they will have the same external behavior under all possible schedules. In such a case, by fixing a schedule, the verification problem of a MapReduce program can be reduced to the verification problem of a sequential program, which is known to be much easier then the verification problem of concurrent programs. One needs to consider only a particular interleaving of input values instead of all of them.

@@

We use two sets of integer reducer benchmarks to evaluate J-ReCoVer. The first set is collected from open repositories such as Github and Bitbuckect. With the help of a search engine~\cite{searchcode} over those repositories, we collected 118 examples. J-ReCoVer is able to correctly analyze all but 3 of them. We also tried all 10 integer reducers descried in verification literature~\cite{ChenHSW15,ChenSW16} and J-ReCoVer can handle all of them.

\subsection*{Related Works}
Reducer commutativity problem can be reduced to program equivalence problem. One creates another program $R'$ that first non-deterministically swaps two consecutive input values and then executes the code of $R$. If $R'$ and $R$ are equivalent, using the fact that all permutations of a list can be obtained by swapping consecutive list elements finitely many times\footnote{Here is an example to produce [3;2;5;1;4] from [1;2;3;4;5] by swapping consecutive elements: [\textbf{1};\textbf{2};3;4;5] $\rightarrow$ [2;\textbf{1};\textbf{3};4;5] $\rightarrow$ [2;3;1;\textbf{4};\textbf{5}]$\rightarrow$ [2;3;\textbf{1};\textbf{5};4]$\rightarrow$ [\textbf{2};\textbf{3};5;1;4]$\rightarrow$ [3;2;5;1;4].}, $R$ can be proved to be commutative. There are a series of research results on program equivalence checking (or closely related topics such as regression verification)~\cite{symdiff,fedyukovich2015automated,sharma2013data,godlin2009regression,fedyukovich2016property,felsing2014automating,KieferKlebanovUlbrich2017,lahiri2013differential,grossman2017verifying,barthe2011relational,KlebanovRuemmerUlbrich2017}. At a high level, checking equivalence of two programs $P$ and $P'$ can be reduced to a sequential verification problem by executing $P'$ after $P$ and at the end checking if they always have the same outputs. It can be made more efficient by finding the right~\emph{synchronization points} and combining the code of $P$ and $P'$ in an interleaved manner. A lot of the research effort in this research direction is in finding good synchronization points. For the case of commutativity analysis, good synchronization points are already given; one can simply use the beginning of the loop as the synchronization point. However, in our experience, if we naively reduce the commutativity problem to equivalence problem and check it in a precise manner, many examples cannot be solved. In this paper, we propose a good over-approximation (c.f. Section~\ref{sec:prover}) of the reducer's behavior. The approximation allows a much more efficient yet precise enough commutativity analysis.

@@

\section{Notations and Definitions}
\label{section:integer-reducers}
We use $\Z$ to denote the set of integers and $[n,m]$ to denote the set of integers $\{k\mid n \leq k\leq m\}$. 
We lift the equality $=$ operation to tuples in the standard way.
Let $S\subseteq [1,k]$, the equality $\{v_1,\ldots,v_k\}=_{S} \{v'_1,\ldots,v'_k\}$ between tuples holds if and only if $\bigwedge_{i\in S} v_i=v'_i$.
For presentation purpose, we describe a highly simplified model to characterize the essence of reducers. The model allows to describe the computation of reducers and investigate their commutativity.

\begin{equation*}
\begin{array}{rclcl}
\multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and Constant}\\
e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
g\in \Grd  & {\triangleq}  & e < e \mid not\ g \mid g \wedge g  \mid \ldots& & \textmd{Guard}\\
s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && \textmd{Command}\\
p\in \Prg  & {\triangleq}  & s.\rloop\{s\}.s&& \textmd{Reducer Program}\\
\end{array}
\end{equation*}

Here $v\in \Var$ and $c \in \Z$ are variables and constant integer values, respectively. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{guard} in $\Grd$ can be a comparison between two expressions or boolean combinations of several guards. A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each branch\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing every branch statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g)\wedge s_1$ or $(not\ g) \wedge s_2$.} . According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. A few examples of reducer programs in this model can be found in Figure~\ref{fig:reducer_example}.

Some reducer uses two loops to compute the output. For example, for calculating standard deviation, one first computes the average of inputs and then uses it to compute the standard deviation. In that case, we can still verify the reducer by first partitioning it into two reducers and then verifying them separately. One reducer computes the average (Figure~\ref{fig:reducer_example}(b)) and the other reducer uses a non-deterministic value to over-approximate the computed average ($a := *$) and then use it in the computation of standard deviation (Figure~\ref{fig:reducer_example}(c)). The reducer in Figure~\ref{fig:reducer_example}(c) ``over-approximates''  all possible output of the standard deviation computation; it computes an output even if $a$ is not the average of inputs. However, the reducer is still commutative and its commutativity verification can still be done.

\begin{figure}
	\begin{minipage}{0.32\textwidth}
		\begin{algorithm}[H]
			$m := \cur$; \;
			\Loop{}{
				$t:=\cur$;\;
				\uIf{ $t > m$}{
					$m := t$ 
				}
			};
			$out(m)$\;
		\end{algorithm}
		\caption*{(a) max}
	\end{minipage}	
	\hspace{-0.5cm}
	\begin{minipage}{0.3\textwidth}
		\begin{algorithm}[H]
			$s := 0;c:=0$; \;
			\Loop{}{
				$s := s+\cur$;\;
				$c := c+1$ 
			};
			$o := s/c$;\;
			$out(o)$\;
		\end{algorithm}
		\caption*{(b) average}
	\end{minipage}	
	\begin{minipage}{0.4\textwidth}
		\begin{algorithm}[H]
			$a:=*;s := 0; c:=0$;\;
			\Loop{}{
				$s := s+(\cur-a)^2$; \;
				$c := c+1$ 
			};
			$o := s/c$;\;
			$out(o)$
		\end{algorithm}
		\caption*{(c) standard deviation}
	\end{minipage}	
	\caption{Example of reducer programs}
	\label{fig:reducer_example}
\end{figure}

@@

\section{Overview of the J-ReCoVer tool}
\label{sec:overview}
%Hadoop, Java;
The input of J-ReCoVer is a reducer program written in Java, which is the most popular programming language used in the Hadoop~\cite{hadoop} MapReduce framework. 
The J-ReCoVer tool has three main components, the \emph{Preprocessor}, the \emph{Prover}, and the \emph{BugFinder}. As the name suggests, the Preprocessor reads as input a reducer program and performs the required preprocessing.  The goal of Prover is to show that a given reducer is commutative and the goal of BugFinder is the opposite. The architecture of J-ReCoVer can be found in Figure~\ref{fig:overview-jrecover}. The user can input a reducer program to J-ReCoVer either through our web-interface or binary installed on their own machine.

@@

\textbf{The Preprocessor component} first compiles a reducer program to bytecode and use the tool Soot~\cite{soot} to further convert it to the so called ``Jimple'' format, which is an intermediate language that is designed to simplift the analysis of Java programs. Under the Hadoop MapReduce framework, the permutation of the input is handled by the scheduler/shuffler component and is affected by things like network latency, which is not controllable by programmers. In order to deal with this issue, we wrote our own dummy Hadoop environment for the reducer as a part of the Preprocessor component. So that the input order of the reducer is now controlled by J-ReCoVer. Finally, we perform in the Preprocessor some simple program transformations tasks to simplify the analysis. For example, we convert all output commands $out(e)$ inside loops to an special assignment (Section~\ref{sec:program_trans1} and~\ref{sec:program_trans2}).

\textbf{The BugFinder component} randomly generated 1500 pairs of lists. The lists in a pair are permutations of each other. A concrete counterexample is reported if inputting any pair of lists from the 1500 generated ones to the reducer generates different results.
Our procedure for generating random pairs is quite naive. We use five different input list of lengths 5, 7, 9 ,11, and 13. For each length, we generate 300 lists and pick uniformly at random one of its permutations. Although the approach is simple, in practice it finds counterexamples in all of our non-commutative benchmarks.

\textbf{The Prover component} reduces the commutativity problem to a SMT formula satisfiability problem. At a high level, we check equivalence between a reducer program and its variant that have two consecutive inputs swapped. We show that this can be reduced to a SMT problem and give it to SMT solver Z3~\cite{z3} for solving. Recall that all permutations of a list can be obtained by swapping consecutive elements finitely many times. In case that Z3 proved the formula unsatisfiable, we know that swapping any two consecutive inputs of the reducer will not change its output. It follows that the reducer outputs the same value for all permutations of the same list of input.
In this case,  J-ReCoVer stops and reports that the reducer is commutative.

If we run the approach in a naive manner, it does not scale to large programs and is very imprecise.  We will introduce two critical optimizations that enables the proposed approach working for real-world examples in Section~\ref{section:optimizations}: a data flow analysis to extract critical variables and an improved symbolic execution algorithm to generate Z3 formulae. More details of the algorithms used in the Preprocessor and the Prover components will be explained in Section~\ref{sec:preprocessor_prover} and the optimizations will be introduced in~\ref{section:optimizations}.

@@

\subsection{The First Program Transformation Task in the Preprocessor}
\label{sec:program_trans1}
J-ReCoVer  first generates a new variable $v_{out_k}$ for the $k$-th output command $out(e)$ inside $s_2$ and replace the $out(e)$ command with an assignment $v_{out_k}:=e$. 
Notice that the value of $v_{out_k}$ stays the same from the point it has been assigned a value. 
If the values of $v_{out_k}$ are the same for all permutations of input lists at the end of $s_2$, we know that the reducer is commutative w.r.t. to all output values produced inside the loop body $s_2$.

\subsection{The Prover: Reducing Commutativity to SMT Solving}
\label{sec:prover}
Below we begin with a simple version where $s_1$ never calls the $\cur$ function. Later we will explain how to handle the general case using program transformation. The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1} \rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$, where $k$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively. The construction of $F$ from $s_2$ can be done in the standard way. We will introduce a more efficient data structure for the construction in Section~\ref{section:optimizations}.

We check if the formula below is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$. 
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu}
\end{equation}

@@

In the example of ``average'', the values of $s$ and $c$ are the same no matter the reducer first reads $n_1$ and then $n_2$ or reads the two values in the swapped order.
So it computes the same value for the input [1;2;3;4;5] and any other input with two consecutive values swaps, e.g., [1;2;\textbf{4};\textbf{3};5].
Since any permutation of a sequence can be obtained by swapping consecutive values finitely many times~\cite{algebra}, this further implies that starting from the same variable valuation, the loop ends with the same valuation for any permutation of input values, e.g., the reducer will output the same value for [1;2;\textbf{3};\textbf{4};5], [1;\textbf{2};\textbf{4};3;5], [\textbf{1};\textbf{4};2;3;5], [4;1;2;3;5], etc. 
Recall that the first program transformation task remembers all output values in variables that will stay unchanged until the end of the loop. The validity of Equation~(\ref{eq:commu}) also implies that the execution of the loop produces the same sequence of outputs.

We explain why that the validity of Equation~(\ref{eq:commu}) is sufficient to prove the commutativity of the reducer $s_1;\rloop\{s_2\};s_3$. Notice that $s_1$ does not read from the input (assumed at the beginning of this subsection). The execution of $s_1$ is deterministic in the sense that from the same initial valuation, it outputs the same sequence of values and ends at the same valuation. For the execution of $\rloop\{s_2\}$, from the validity of Equation~(\ref{eq:commu}), any permutation of input values leads the reducer to the same final valuation and the same sequence of outputs at the end of the loop. 
Before executing the commands after the loop $s_3$, all inputs have been read (the loop exit condition) already, similar the case of $s_1$, the execution of $s_3$ is deterministic for all permutations of inputs and outputs the same sequence of values.

@@

In real-world examples, it is often that the $s_1$ part of the reducer also reads from the input. In the example in Figure~\ref{fig:reducer_max} (a), the reducer program remembers the first input value in the variable $m$, increases its value by 10, and then updates its value to bigger ones if any occurs in the loop. The corresponding Equation~(\ref{eq:commu}) is valid in this case. However, the reducer is not commutative. A counterexample can be easily found. With the list $[1,2,3,4,5]$ and its permutation $[5,4,3,2,1]$ as the inputs, the reducer respectively outputs $11$ and $15$.

We handle the problem again by program transformation. To handle the example in Figure~\ref{fig:reducer_max} (a), we move the prefix $s_1$ into the loop body and use a new variable $s$ to force that $s_1$ is always executed before the original loop body. The result after the transformation is demonstrated in Figure~\ref{fig:reducer_max} (b). Observe that the new reducer program over-approximates all possible inputs/outputs of the original one.
It also satisfies all assumptions required for the Prover component: (a) The commands before entering the loop does not have any $\cur$ expression and (II) each branch (recall that ``branch'' is formally defined in Section~\ref{section:integer-reducers}) of the loop body has at most one $\cur$. So we can apply Equation~(\ref{eq:commu}) to prove its commutativity.
Since the new reducer has all the possible outputs of the original one, the new reducer being commutative implies that the original one is also commutative.

\begin{figure}[htb]
	\centering
	\begin{minipage}{0.8\textwidth}
	\begin{algorithm}[H]
		$s'_0;s:=*;$\;
		\Loop{}{
			\lIf{$s=1$}{ $x_{i_1} := \cur; s'_1; s:=2$}
			\lElseIf{ $s =2$}{ $x_{i_2} := \cur; s'_2; s:=3$}
			\lElseIf{ $s = 3$}{$\ldots$}
			\lElseIf{ $s = m$}{$x_{i_m} := \cur; s'_m; s:=m+1$}
			\lElse{ $s_2$}
		};$s_3$\;
	\end{algorithm}
	\end{minipage}
	\caption{The second transformation task}
	\label{fig:general_program_transformation}
\end{figure}

For the general case, we first rewrite the $s_1$ part of the reducer $s_1;\rloop\{s_2\};s_3$ to the form 
$$s'_0;x_{i_1}{:=} \cur;s'_1.\ldots x_{i_{m}}{:=}\cur;s'_m, $$ where $\cur$ never occur in $s'_0,s'_1,\ldots, s'_m$. Then we produce the reducer in Figure~\ref{fig:general_program_transformation}, which is an over-approximation of the original reducer. For the case that $s=1$ at the beginning, its output sequence is the same as the original reducer. When $s\neq 1$, it may produce outputs that are not possible by the original reducer. Observe that the reducer in Figure~\ref{fig:general_program_transformation} satisfies all assumptions required for the Prover: each branch has at most one $\cur$. Hence the Equation~(\ref{eq:commu}) can then be applied to prove its commutativity.

\section{Optimizations}
\label{section:optimizations}
In this section we explain how data flow analysis are used in J-Recover to improve the precision of commutativity checking and the algorithm we used to construct the $F$ function in Equation~(\ref{eq:commu}).

\subsection{Improving the precision of commutativity verification using data flow analysis}
\label{section:opt1}
We find that in many cases, Equation~(\ref{eq:commu}) is a condition that is too strong for the reducer to be commutative. For instance, consider the example in Figure~\ref{fig:reducer_opt}, in the loop body, the input is first stored in a variable $t$ and then added to $s$. In this case, the $F$ function returns the updated values of both $s$ and $t$ after the execution of loop body. Observe that 
$F(c_1, F(c_2,s,t) = (c_1+c_2, c_1)$ and $F(c_2, F(c_1,s,t) = (c_1+c_2, c_2)$. It follows that the corresponding Equation~(\ref{eq:commu}) is invalid. The second component of the returned tuple, which corresponds to the value of $t$ after executing the loop body twice, are $c_1$ and $c_2$, respectively. However, in this case, the value of $t$ is not important to the output of this reducer.
